---
title: "Assignment 01 - selected keys"
output:
  html_document:
    df_print: paged
---

####2b. Getting last three elements
```{r}
health.expenditures <- c(864, 320, 239, 882, 472, 72, 60, 607, 3984, 4668, 47, 180)
# method 1
l <- length(health.expenditures)
health.expenditures[(l-2):l]

# method 2
tail(health.expenditures, 3)

#method 3
rev(health.expenditures)[1:3]
rev(rev(health.expenditures)[1:3]) #if the order matters
```

####6a. Indexing phone book
```{r}
#create a phone book
john <- list(fname = "John", lname = "Smithee", age = 24,
             phone = "919-555-1234")
jane <- list(fname = "Jane", lname = "Doe", age = 39,
             phone = "919-555-4321")
jack <- list(fname = "Jack", lname = "Indabox", age = 13,
             phone = NA)
jill <- list(fname = "Jill", lname = "Bopeep", age = 18,
             phone = "919-555-1243")

phone.book <- list(john, jane, jack, jill)

#index to the second person
phone.book[[2]]

#index to `fname` and `lname` of the second person
phone.book[[2]][c("fname","lname")]
typeof(phone.book[[2]][c("fname","lname")])